---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = '100%'
)
```

# rsinaica

[![Travis-CI Build Status](https://travis-ci.org/diegovalle/rsinaica.svg?branch=master)](https://travis-ci.org/diegovalle/rsinaica) 
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/p281myk561l2kxgt?svg=true)](https://ci.appveyor.com/project/diegovalle/rsinaica/branch/master) 
[![Coverage Status](https://img.shields.io/codecov/c/github/diegovalle/rsinaica/master.svg)](https://codecov.io/github/diegovalle/rsinaica?branch=master)
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-red.svg)](https://www.tidyverse.org/lifecycle/#experitmental) 
[![CRAN\_Status\_Badge](http://www.r-pkg.org/badges/version-ago/rsinaica?color=red)]()

Ready-made functions for downloading air quality data from the Mexican National 
Air Quality Information System (SINAICA).

## Installation

rsinaica is not currently available from CRAN, but you can install the development version from github with:

```r
if (!require(devtools)) {
    install.packages("devtools")
}
devtools::install_github('diegovalle/rsinaica')
```


<!--ou can install the released version of rsinaica from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("rsinaica")
```-->

## Example

Suppose you wanted to download pollution data from the _Centro_ station in
Guadalajara. First we load the necessary packages and look up the numeric code
for the station in the `stations_sinaica` data.frame:

```{r example, warning=FALSE,message=FALSE}
## Auto-install required R packages
packs <- c("ggplot2", "maps", "mapproj")
success <- suppressWarnings(sapply(packs, require, character.only = TRUE))
if (length(names(success)[!success])) {
  install.packages(names(success)[!success])
  sapply(names(success)[!success], require, character.only = TRUE)
}
library("rsinaica")

knitr::kable(stations_sinaica[which(stations_sinaica$station_name == "Centro"), 1:6])
```

It looks like there are three stations named Centro, the one we are looking for is the one in Guadalajara with a numeric code of 102 (station_id). The ```stations_sinaica``` data.frame also includes the latitude and longitude of all the measuring stations in Mexico (including some that have never reported any data!).

```{r warning=FALSE}
mx <- map_data("world", "Mexico")
stations_sinaica$color <- "Others"
stations_sinaica$color[stations_sinaica$station_id == 102] <- "Centro (102)"
ggplot(stations_sinaica[order(stations_sinaica$color, decreasing = TRUE),], aes(lon, lat)) + 
  geom_polygon(data = mx, aes(x= long, y = lat, group = group)) +
  geom_point(alpha = .9, size = 3, aes(fill = color), shape = 21) + 
  scale_fill_discrete("station") +
  ggtitle("Air quality measuring stations in Mexico") +
  coord_map() + 
  theme_void()
```

Then we query the dates during which the station has been in operation:

```{r}
get_station_dates(102)
```

It's currently reporting data, and has been doing so since 1997. We can also query which type of parameters (pollution, wind, solar radiation, etc) the station has sensors for. Note that the package also includes a ```parameters```
data.frame with the complete set of supported parameters, but not all stations support all of them.

```{r}
cen_params <- get_station_parameters(102)
knitr::kable(cen_params)
```

Finally, we can download and plot particulate matter with a diameter between 2.5 and 10 micrometers (μm) (PM<sub>10</sub>) for the month of January

```{r}
# Download all PM10 data for January 2018
df <-  sinaica_bystation(102, "PM10", "2018-01-01", "1 month", "Crude")

ggplot(df, aes(hour, value, group = date)) +
  geom_line(alpha=.9) +
  ggtitle(expression(paste(PM[10],
                           " pollution during January 2018 in Centro, Guadalajara, by hour"))) +
  xlab("hour") +
  ylab(expression(paste(mu,"g/", m^3))) +
  theme_bw()
```
