---
title: "Comparing rsinaica to aire.zmvm"
author: "Diego Valle-Jones"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

First, we load the packages

```{r}
## Auto-install required R packages
packs <- c("dplyr", "ggplot2", "aire.zmvm")
success <- suppressWarnings(sapply(packs, require, character.only = TRUE))
if (length(names(success)[!success])) {
  install.packages(names(success)[!success])
  sapply(names(success)[!success], require, character.only = TRUE)
}
## Not yet on cran
library("rsinaica")
```

Then we download the data and make sure that both the values from aire.zmvm and rsinaica are in ppb
(parts per billion), since rsinaica returns values in ppm (parts per million) we multiply by 1,000. We also make sure to set the datetime to GMT+6 (Mexico City time with no DST)

```{r}
df_aire <- get_station_month_data("HORARIOS", "O3", 2018, 4) %>%
  filter(station_code == "XAL") %>%
  mutate(value_aire.zmvm = value)
df_aire$datetime <-  as.POSIXct(
  strptime(paste0(df_aire$date, " ", df_aire$hour),
           "%Y-%m-%d %H", tz = "Etc/GMT+6")
)
df_aire <- df_aire[, c("datetime", "value_aire.zmvm")]


## values from sinaica are in ppb and those from aire.cdmx in ppm
## Xalostoc station has an rsinaica station_id of 271
df_sinaica <- sinaica_station_data(271, "O3", "2018-03-01", "2018-03-30") %>%
  mutate(value = value * 1000) %>%
  mutate(value_sinaica = value) %>%
  mutate(date = as.Date(date))
df_sinaica$datetime <-  as.POSIXct(
  strptime(paste0(df_sinaica$date, " ", df_sinaica$hour),
           "%Y-%m-%d %H", tz = "Etc/GMT+6")
)
df_sinaica <- df_sinaica[, c("datetime", "value_sinaica")]

df <- inner_join(df_aire, df_sinaica, by = c("datetime" = "datetime"))
```

This is what the data looks like:

```{r echo=FALSE}
knitr::kable(df[5:10, ])
```

We can visually compare them with a plot. The values are extremely similar, but the aire.zmvm data is a little bit more complete since it comes directly from the source.

```{r}
ggplot(df_aire, aes(datetime, value, color = "aire.zmvm")) +
  geom_line(size = 1.5, alpha = .4) +
  geom_line(data = df_sinaica, aes(datetime, value, color = "rsinaica"),
            alpha = .4, size = 1.5) +
  xlab("datetime (GMT+6)") +
  ylab("ppb") +
  scale_color_discrete("package") +
  ggtitle(expression(paste("Hourly ", O[3], " pollution values during March"))) +
  theme_bw()
```
