---
title: "Most PM2.5 Polluted City in Mexico - 2017"
author: "Diego Valle-Jones"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


We can use `rsinaica` to find out which city is the most PM<sub>25</sub>-polluted in all of Mexico.

First, we load the packages:

```{r packages,warning=FALSE,message=FALSE}
## Auto-install required R packages
packs <- c("dplyr", "ggplot2", "gghighlight", "lubridate", "anomalize", 
           "aire.zmvm", "tidyr", "zoo", "plotly")
success <- suppressWarnings(sapply(packs, require, character.only = TRUE))
if (length(names(success)[!success])) {
  install.packages(names(success)[!success])
  sapply(names(success)[!success], require, character.only = TRUE)
}
## Not yet on cran
library("rsinaica")
```

## Download

Then we download the data for the whole year of 2017 using the `sinaica_param_data` function. Since the maximum data range we can download is 1 month, we have to use a little mapply magic to download the entire year.

```{r download_pm25, cache=TRUE}
# Download all PM25 pollution data in 2017
pm25_2017 <- bind_rows(
  mapply(sinaica_param_data,
         "PM2.5",
         seq(as.Date("2017-01-01"), as.Date("2017-12-01"), by = "month"),
         seq(as.Date("2017-02-01"), as.Date("2018-01-01"), by = "month") - 1,
         SIMPLIFY = FALSE)
)
       
```

I thought there would be a few cities that collected PM<sub>25</sub> data manually (they collect it through a filter and send it to be weighted to an external lab, sometimes in another country). But no air quality station collected manual PM<sub>2.5</sub> in 2017.

```{r}
bind_rows(
  mapply(sinaica_param_data,
         "PM2.5",
         seq(as.Date("2017-01-01"), as.Date("2017-12-01"), by = "month"),
         seq(as.Date("2017-02-01"), as.Date("2018-01-01"), by = "month") - 1,
         "Manual",
         SIMPLIFY = FALSE)
)
```

This is what the data looks like:

```{r}
knitr::kable(head(pm25_2017))
```

## Cleanup

Once we've downloaded the data we filter values below 1 µg/m³ since they're probably calibration errors. And we only include stations that reported for more than 80% of days (292). We also have to take into account that PM<sub>25</sub> data is measured as a 24 hour average.

```{r plot}
# pm25_2017[which(pm25_2017$value_actual != pm25_2017$value_original),]
# pm25_2017[which(!is.na(pm25_2017$date_validated)),]

## filter stations that didn't report at least 47 weeks of the year
df_filtered <- pm25_2017 %>%
  #filter(value < 250) %>%
  mutate(value = if_else(value < 1, NA_real_, value)) %>%
  group_by(network_name) %>%
  filter(!is.na(value)) %>%
  mutate(nweeks = n_distinct(week(date))) %>%
  filter(nweeks >= 47) %>%
  select(-nweeks) %>%
  ungroup()



df_max <- df_filtered %>%
  complete(station_id,
           hour = 0:23,
           date = as.character(seq(as.Date("2017-01-01"), as.Date("2017-12-31"), by = "day"))) %>%
  group_by(station_id, network_name) %>%
  arrange(station_id, date, hour) %>%
  mutate(roll24 = rollapply(value, 24, mean, na.rm = TRUE, partial = 18, 
                            fill = NA, align = "right")) %>%
  ungroup() %>%
  #summarise(mean = mean(value, na.rm = TRUE)) %>%
  group_by(date, network_name) %>%
  summarise(max = max(roll24, na.rm = TRUE)) %>%
  ungroup() %>%
  add_count(network_name) %>%
  ## Only include stations that reported for more than 80% of days (292)
  filter(n >= (365*.8))  %>%
  select(-n) %>%
  filter(is.finite(max)) %>%
  arrange(date)
```

When plotting the daily 24 hour average maximums we can see that there are still some obvious errors in the data.

```{r fig.width = 8, fig.height = 18}
ggplot(df_max, aes(as.Date(date), max)) +
  geom_line(size = .3, color = "black") +
  facet_wrap(~ network_name, ncol = 3) +
  ggtitle(expression(paste("Maximum daily ", PM[2.5], " concentration by network"))) +
  xlab("date") +
  ylab(expression(paste("daily maximum 24 average of ", PM[2.5], 
                        " (", mu,"g/", m^3, ")"))) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=60,hjust=1))
```

It looks like we can safely remove values above 200 µg/m³ and get rid of the Aguascalientes, Irapuato and Monclova networks. 

## Anomalies

We can also use the `anomalize` package to detect extreme values, but actually figuring out if they are errors is a little bit more tricky since fires can temporarily spike PM<sub>25</sub> levels as often happens during the winter holidays when people burn trash and tires, and set off fireworks. I've opted *not* to remove them, because of the spikes around new year. Since I'm interested in the average of the whole year, these PM<sub>25</sub> outliers are unlikely to have a substantial effect on the rankings.

```{r fig.width = 8, fig.height = 18}
df_max <- filter(df_max, !network_name %in% 
                   c("Irapuato", "Monclova", "Aguascalientes")) %>%
  filter(max <= 200)

df_max <- df_max %>%
  ungroup() %>%
  group_by(network_name) %>%
  mutate(date = as.Date(date)) %>%
  time_decompose(max, method = "stl") %>%
  anomalize(remainder, method = "iqr", alpha = 0.01) %>%
  time_recompose()


# Anomaly Visualization
df_max %>% plot_anomalies(time_recomposed = TRUE, ncol = 3, alpha_dots = 0.25) +
  labs(title = "Tidyverse Anomalies", subtitle = "STL + GESD Methods") 

## Don not remove outliers, if you look at the chart some values around new year
## are detected as spikes, but it's people setting off fireworks and burning trash
## df_max <- filter(df_max, anomaly != "Yes")
```

It was decided to remove Irapuato from the rankings and keep Tepic because PM<sub>2.5</sub> values can spike because of fires. Normally I would have removed Mexicali since it didn't report any correct data during the winter season, but its levels of PM<sub>2.5</sub> where so high that I decided to keep it.

## Most PM<sub>25</sub>-Polluted City

And here is the most PM<sub>25</sub>-polluted city in Mexico: <b>Toluca</b>

```{r}

gghighlight_line(df_max, aes(as.Date(date), observed, 
                              group = network_name, color = network_name),
                 mean(observed, na.rm = TRUE), max_highlight = 1) +
  theme_bw() +
  ggtitle(expression(paste("Pollution measuring network with the highest ", 
                           PM[2.5], " pollution values in 2017")),
          subtitle = "Based on the mean of the highest 24-hour rolling average daily maximums. Source: SINAICA") +
  xlab("date") +
  ylab(expression(paste(PM[2.5], " ", mu,"g/", m^3)))
```

## Top polluted cities

```{r}
knitr::kable(
  df_max %>%
    group_by(network_name) %>%
    summarise(mean = mean(observed, na.rm = TRUE)) %>%
    arrange(-mean) %>%
    head(10)
  )
```

The top 3 most PM<sub>25</sub>-polluted cities in Mexico. Note that Mexicali didn't report any data during the winter, which is the season with the highest levels of PM<sub>2.5</sub> pollution, it could probably have ranked higher if they bothered to keep their instruments working.

```{r plotly, warning=FALSE, message=FALSE, fig.width = 10, fig.height = 4}
top3 <- filter(df_max, network_name %in% c("Toluca",
                                           "Cuernavaca",
                                           "Guadalajara"))[, 1:3]

top3 <- spread(top3, network_name, observed)
top3$date <- as.Date(as.Date(top3$date))
#top3$Toluca <- na.spline(top3$Toluca)
#top3$Guadalajara <- na.spline(top3$Guadalajara)
#top3$Mexicali <- na.spline(top3$Mexicali)
x <- list(
  title = "date"
)
y <- list(
  title = "daily maximum 24 average of PM2.5"
)

plot_ly(as.data.frame(top3), x = ~date, y = ~`Toluca`, name = "Toluca" ,
        type = 'scatter', mode = 'lines', line = list(color = '#e41a1c'), width = .5) %>%
  add_trace(y = ~Guadalajara, name = 'Guadalajara', mode = 'lines', line = list(color = '#377eb8'), width = .5)  %>%
  add_trace(y = ~Cuernavaca, name = 'Cuernavaca', mode = 'lines', line = list(color = '#4daf4a'), width = .5) %>%
  layout(title = "Top 3 most PM2.5-polluted cities in Mexico", xaxis = x, yaxis = y)
```

## Days with bad air quality

Number of days with bad air quality (Índice IMECA MALO)

```{r}
df_max %>%
  group_by(network_name) %>%
  filter(observed > 45.05) %>%
  summarise(count = n()) %>%
  arrange(-count) %>%
  head() %>%
  knitr::kable()
```

Number of days with very bad air quality (Índice IMECA MUY MALO).

```{r}
df_max %>%
  group_by(network_name) %>%
  filter(observed > 97.45) %>%
  summarise(count = n()) %>%
  arrange(-count) %>%
  head() %>%
  knitr::kable()

```
