---
title: "Long-term pollution trends"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

With `rsinaica` we can plot the long term trends in pollution for different cities. First, we load the packages

```{r packages, warning=FALSE,message=FALSE}
## Auto-install required R packages
packs <- c("dplyr", "ggplot2", "lubridate", "scales")
success <- suppressWarnings(sapply(packs, require, character.only = TRUE))
if (length(names(success)[!success])) {
  install.packages(names(success)[!success])
  sapply(names(success)[!success], require, character.only = TRUE)
}
## Not yet on cran
library("rsinaica")
```

We create a function for downloading single pollutant data for all the stations that form a network

```{r function, warning=FALSE}
get_network_data <- function(network_name, pollutant, remove_extremes = TRUE) {
  ## Download data for all Aguascalientes stations for a single month
  get_month <- function(start_date, end_date, net){
    bind_rows(
      lapply(stations_sinaica$station_id[stations_sinaica$network_name %in% net],
             sinaica_station_data, pollutant, start_date, end_date, "Crude",
             remove_extremes = remove_extremes)
    )
  }
  ## Download data for 2017, by month
  df <- bind_rows(
    mapply(get_month,
           seq(as.Date("1997-01-01"), as.Date("2017-12-01"), by = "month"),
           seq(as.Date("1997-02-01"), as.Date("2018-01-01"), by = "month") - 1,
           network_name, 
           SIMPLIFY = FALSE)
  )
  
  df$datetime <-  with_tz(as.POSIXct(paste0(df$date, " ", df$hour, ":00"), 
                                     tz = "Etc/GMT+6"),
                          tz = "America/Mexico_City")
  df
}
```

Now we are ready to plot different cities. When requesting ozone values with the `sinaica_station_data` function, there is an option to automatically clean the data and set values above .2 ppm to NA, since Guadalajara tends to have quite high ozone levels it is probably a good idea to disable this functionality.

```{r guadalajara, cache=TRUE, warning=FALSE}
get_network_data("Guadalajara", "O3", remove_extremes = FALSE) %>%
  ## remove ozone values above .3 cause Guadalajara air is extremely dirty
  filter(value < .3 & value >= 0) %>%
  ggplot(aes(as.Date(date), value)) +
  geom_point(alpha = .01, size = .3) +
  geom_smooth(method = 'gam', formula = y ~ s(x)) + 
  ggtitle("Daily ozone values reported by all Guadalajara stations") +
  xlab("date") +
  ylab("ppm") +
  coord_cartesian(ylim = c(0, 0.3)) +
  theme_bw()
```

```{r toluca, cache=TRUE, warning=FALSE}
get_network_data("Toluca", "PM2.5", remove_extremes = FALSE) %>%
  filter(value <= 750) %>%
  ggplot(aes(as.Date(date), value)) +
  geom_point(alpha = .01, size = .3) +
  geom_smooth(method = 'gam', formula = y ~ s(x)) + 
  ggtitle(expression(paste("Daily ", PM[2.5], 
                           " values reported by all Toluca stations"))) +
  xlab("date") +
  ylab(expression(paste(mu,"g/", m^3))) +
  coord_cartesian(ylim = c(0, 210)) +
  theme_bw()
```

```{r pachuca, cache=TRUE, warning=FALSE}
get_network_data("Pachuca", "PM2.5", remove_extremes = FALSE) %>%
  filter(value <= 600) %>%
  ggplot(aes(as.Date(date), value)) +
  geom_point(alpha = .01, size = .3) +
  geom_smooth(method = 'gam', formula = y ~ s(x)) + 
  ggtitle(expression(paste("Daily ", PM[2.5], 
                           " values reported by all Pachuca stations"))) +
  xlab("date") +
  ylab(expression(paste(mu,"g/", m^3))) +
  coord_cartesian(ylim = c(0, 150)) +
  theme_bw()
```


```{r monterrey, cache=TRUE, warning=FALSE}
get_network_data("Monterrey", "PM10", remove_extremes = FALSE) %>%
  filter(value <= 1000) %>%
  ggplot(aes(as.Date(date), value)) +
  geom_point(alpha = .01, size = .3) +
  geom_smooth(method = 'gam', formula = y ~ s(x)) + 
  ggtitle(expression(paste("Daily ", PM[10], 
                           " values reported by all Monterrey stations"))) +
  xlab("date") +
  ylab(expression(paste(mu,"g/", m^3))) +
  coord_cartesian(ylim=c(0, 500)) +
  theme_bw()
```

